A Large-Scale Study of Polymorphism in R

### 2019-04-06 paper/main.tex

Figure~\ref{fig:charclos} shows an example of a function with character and
function polymorphism.  At a high level, this function
calculates point estimates for an {\tt angmcmc} object (specific to some
packages).  The caller specifies {\tt fn}, which is either a function {\it or} the name of a function which will be evaluated on the object samples to estimate parameters.
If {\tt fn} is a string, the function lookup is performed by \code{match.fun}.

@@

More generally, there is an existing literature on analyzing usage patterns of language features.  
For instance, the dynamic features of Smalltalk were analyzed in some work~\cite{callau2011howdevelopers}, which aimed to see how often the highly dynamic capabilities of Smalltalk were used and in which kinds of projects.
Other work~\cite{pldi10a} explores how the dynamic features of JavaScript are used.
An interesting finding of theirs is that

Other analysis work~\cite{milojkovic2017duck} ... .

The type system we propose in this work could be incorporated into R as a {\it gradual type system}.
Gradual type systems~\cite{SiekTaha06} aim to bridge the gap between dynamic and static typing by allowing programmers to gradually add type annotations to their programs.
Gradually-typed languages are becoming increasingly popular, with languages such as TypeScript~\cite{typescript13} becoming more mainstream.
...

>>>>

Figure~\ref{fig:charclos} shows an example of a function with character and
function polymorphism.  At a high level, this function
calculates point estimates for some object. The caller specifies {\tt fn}, which is either a function {\it or} the name of a function which will be evaluated on the object samples to estimate parameters.
If {\tt fn} is a string, the function lookup is performed by \code{match.fun}.

@@

More generally, there is an existing literature on analyzing usage patterns of language features.  
For instance, the dynamic features of Smalltalk were analyzed in some work~\cite{callau2011howdevelopers}, which aimed to see how often the highly dynamic capabilities of Smalltalk were used and in which kinds of projects.
Other work~\cite{pldi10a} explores how the dynamic features of JavaScript are used.
An interesting finding of theirs is that only 81\% of calls are monomorphic, which is very similar to what we observed in R.
Other analysis work~\cite{milojkovic2017duck} looks into the usage patterns of duck typing.

The type system we propose in this work could be incorporated into R as a {\it gradual type system}.
Gradual type systems~\cite{SiekTaha06} aim to bridge the gap between dynamic and static typing by allowing programmers to gradually add type annotations to their programs.
Gradually-typed languages are becoming increasingly popular, with languages such as TypeScript~\cite{typescript13} becoming more mainstream.
An eventual implementation of a type system for R will surely be inspired by this work.
